---
title: "2 - Preprocess your data with recipes"
author: "by Craig Slinkman"
date: "10/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require("knitr")
opts_knit$set(root.dir <- 
                file.path( "D:/R-Projects/MyTutorials",
                            "Buillding models in the tidverse" ))
```

# Preprocess your data with recipes

## Tidymodels packages: broom, parsnip, recipes, workflows, yardstick
```{r}

```

This is Craig Slinkman's implementation of the tutorial of how to use the tidyverse to carry out systematic model building based on the online tutorial [Preprocess your data with recipes](https://www.tidymodels.org/start/models/).  My implementation will be based on my R-coding style.  Any errors are, of course,  my fault. 


## Loading the required pacakges  

```{r load_packages}
require( tidyverse )                    # I live in the tidyverse.
require( tidymodels )                   # The object of this tutorial.
```

